Parallel Algorithms for Monte Carlo Linear Solvers

### 2015-03-10 doc/siam_cse_2015/siam_cse_2015.tex
adding siam cse 2015 talk
updating first half of cse talk

%%---------------------------------------------------------------------------%%
\begin{frame}{Monte Carlo Synthetic Acceleration}
  \begin{itemize}
  \item Devised by Evans and Mosher in the 2000's as an acceleration
    scheme for radiation diffusion problems (LANL)
    \vfill
  \item Can be abstracted as a general linear solver with Monte Carlo as a
    preconditioner, $M_{MC}$
    \vfill
  \item Combine with Richardson iteration as a ``smoother'' in between
    Monte Carlo steps:
    \begin{align*}
      \mathbf{r}^k &= \mathbf{b} - \mathbf{Ax}^k \\
      \mathbf{x}^{k+1/2} &= \mathbf{x}^k + \mathbf{r}^k \\
      \mathbf{r}^{k+1/2} &= \mathbf{b} - \mathbf{Ax}^{k+1/2} \\
      \mathbf{\delta} &= \mathbf{M_{MC}} \mathbf{r}^{k+1/2} \\
      \mathbf{x}^{k+1} &= \mathbf{x}^{k+1/2} + \delta
    \end{align*}
  \end{itemize}
\end{frame}

@@

  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item Each parallel process owns a piece of the domain (linear
        system)
        \bigskip
      \item Random walks must be transported between adjacent domains
        through parallel communication
        \bigskip
      \item Domain decomposition determined by the input system
        \bigskip
      \item Load balancing straightforward for the forward method
      \end{itemize}
    \end{column}

@@

      \small
      \begin{itemize}
      \item General extension of the Milagro algorithm (LANL)
      \item Asynchronous nearest neighbor communication of histories
      \item System-tunable communciation parameters of buffer size and check
        frequency (performance impact)
      \item Need an asynchronous strategy for exiting the transport loop
        without a collective
      \end{itemize}

      \vspace{-0.25in}
      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.45}{ \input{domain_to_domain.pdftex_t} }
        \end{center}
      \end{figure}

@@

  Different batches of Monte Carlo samplescan be combined in summation via
  superposition if they have different random number streams. For two
  different batches:

>>>>

%%---------------------------------------------------------------------------%%
\begin{frame}{Monte Carlo Synthetic Acceleration}
  \begin{itemize}
  \item Devised by Evans and Mosher in the 2000's as an acceleration
    scheme for radiation diffusion problems (LANL)
    \vfill
  \item Can be abstracted as a general linear solver with Monte Carlo as a
    preconditioner, $\mathbf{M_{MC}}$
    \vfill
  \item Combine with Richardson iteration as a ``smoother'' in between
    Monte Carlo steps:
    \begin{align*}
      \mathbf{r}^k &= \mathbf{b} - \mathbf{Ax}^k \\
      \mathbf{x}^{k+1/2} &= \mathbf{x}^k + \mathbf{r}^k \\
      \mathbf{r}^{k+1/2} &= \mathbf{b} - \mathbf{Ax}^{k+1/2} \\
      \mathbf{\delta} &= \mathbf{M_{MC}} \mathbf{r}^{k+1/2} \\
      \mathbf{x}^{k+1} &= \mathbf{x}^{k+1/2} + \delta
    \end{align*}
  \end{itemize}
\end{frame}

@@

  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item Each parallel process owns a piece of the domain (linear
        system)
        \bigskip
      \item Random walks must be transported between adjacent domains
        through parallel communication
        \bigskip
      \item Domain decomposition determined by the input system
        \bigskip
      \item Load balancing not yet addressed
      \end{itemize}
    \end{column}

@@

    \begin{column}{0.5\textwidth}
      \vspace{-0.1in}
      \small
      \begin{itemize}
      \item General extension of the Milagro algorithm (LANL)
      \item Asynchronous nearest neighbor communication of histories
      \item System-tunable communciation parameters of buffer size and check
        frequency (performance impact)
      \item Need an asynchronous strategy for exiting the transport loop
        without a collective (running sum)
      \end{itemize}

      \vspace{-0.1in}
      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.45}{ \input{domain_to_domain.pdftex_t} }
        \end{center}
      \end{figure}

@@

  Different batches of Monte Carlo samples can be combined in summation via
  superposition if they have different random number streams. For two
  different batches:
